---
title: "2 Exploratory Data Analysis"
author: "Xuelin Zhu"
date: "`r Sys.Date()`"
output: html_document
---
```{r}
load('data/data0.RData')
library(ggplot2)
library(ggcorrplot)
theme_set(theme_classic())
```

This Rmarkdown document is for exploratory data analysis, including descriptive statistics, some plots and correlation. I hope by doing this I can get a better understanding of the data for the modeling.

### 1: Cancer and age
```{r}
tempDf = data.frame(resp = data0[,1], age = data0[,2], gender = data0[,3])
tempDf$resp = factor(tempDf$resp, levels = 0:1, labels = c('Normal', 'Tumor'))
tempDf$gender = factor(tempDf$gender, levels = 0:1, labels = c('Female', 'Male'))

ggplot(tempDf, aes(resp, age)) + 
  geom_boxplot(varwidth=T, fill="plum") + 
  labs(title="Box plot", 
       subtitle="Grouped current age",
       x="Diagnoses",
       y="Current Age")
```

### 2: Cancer and gender
```{r}
ggplot(tempDf, aes(resp)) +
  geom_bar(aes(fill=gender), width = 0.5) + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6)) +
  labs(title="Categorywise Bar Chart", 
       subtitle="Difference between gender", 
       x='Diagnosis',
       y='# of Patients')
```

### 3: Cancer, age and gender
```{r}
ggplot(tempDf, aes(age)) +
  geom_histogram(aes(fill=gender),
                 bins=20, 
                 col="black",
                 size=.1) +  # change binwidth
  labs(title="Age Histogram", 
       subtitle="Difference between gender and age",
       x='Age',
       y='# of Patients')  
```

### 4. Check the contingency table of gender and cancer
```{r}
chisq.test(table(tempDf[,c(1,3)]))
```

### 5. Gene Correlation
Gene data are high dimensional, so I calculated the correlation of first 2000 gene in the next chunk. This procedure takes time, so I have saved its result in the file ```~/data/cor.Rdata``` so that you can skip the next chunk.
```{r, eval=FALSE}
cor = cor(data0[,4:2000])
save(cor, file='data/cor.Rdata')
```

If skipping the last chunk, we can directly read it from the data folder. It is found that many genes are highly correlated.
```{r}
load('data/cor.Rdata')
which(abs(cor)>0.9&cor<1, arr.ind = T)

largeCor = c(256,284,293,307,418,568,661,880,922,925,948,972,999,1139,1230,1440,
             1454,1477,1592,1598,1759,1810,1887,1895)

corr = round(cor[largeCor,largeCor])
rm(largeCor,cor)

ggcorrplot(corr, hc.order = TRUE, 
           type = "lower", 
           lab = TRUE, 
           lab_size = 3, 
           method="square", 
           colors = c("springgreen3", "white", "tomato2"), 
           ggtheme=theme_bw) + 
labs(title="Gene Expression Correlation", 
     subtitle="First 1000 Gene Illustration")
```













