---
title: "5-Modeling"
author: "Xuelin Zhu"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
library(glmnet)
load('data/2-pcreg/PCA.Rdata')
load('data/2-pcreg/trainPcData.Rdata')
load('data/2-pcreg/testPcData.Rdata')
```

### Logistic Regression with PCA
We first check how much variation could be explained by first few pcs.
```{r}
summary(PCA)
```
#### Modeling
First 100 pcs could explain 90% of total variation. So, we use the 100 pcs to fit a logistic regression.
```{r}
logitReg = glm(resp~., data=trainPcData[,1:101], family=binomial())
```

#### Validation
From the warning, it says the MLE still does not converge. Let's check it prediction accuracy.
```{r}
predProb = predict(logitReg, testPcData, type='response')
predResult = predProb>0.5

table(testPcData[,1]==1, predResult, dnn=c('Actual','Predicted'))
```
The behavior is not good. I guess this may due to over-fitting.

#### Auto Selection by AIC
To avoid over-fitting, I did an auto selection by AIC. The next chunk takes a long time, so I have saved its result in ```~/data/2-pcreg/logitRegRed.RData``` so that you can skip it.
```{r, eval=FALSE}
logitRegRed = step(logitReg)
save(logitRegRed, file = 'data/2-pcreg/logitRegRed.RData')
```

Checking the prediction accuracy of reduced logistic model
```{r}
load('data/2-pcreg/logitRegRed.RData')

predProb = predict(logitRegRed, testPcData, type='response')
predResult = predProb>0.5

table(testPcData[,1]==1, predResult, dnn=c('Actual','Predicted'))
```

### Logistic Regression with PCA and Ridge
Let's use Ridge instead of AIC to select significant pcs.
```{r}
trainX = as.matrix(trainPcData[,-1])

mycv = cv.glmnet(trainX, trainPcData[,1], nfolds=10, standardize=TRUE)
plot(mycv);mycv$lambda.min
logitRidge = glmnet(trainX, trainPcData[,1], intercept=TRUE, lambda=mycv$lambda.min); rm(mycv)

testX = as.matrix(testPcData[,-1])

ridgePredResult = predict(logitRidge, testX, type='response')
ridgePredResult = ridgePredResult>0.5

table(testPcData[,1]==1, ridgePredResult, dnn=c('Actual','Predicted'))

save(logitRidge, file='data/3-shrinkage/logitRidge.Rdata')
```

### Logistic Regression with PCA and LASSO
```{r}
mycv = cv.glmnet(trainX, trainPcData[,1], nfolds=10, standardize=TRUE, family='multinomial')
plot(mycv);mycv$lambda.min
logitLasso = glmnet(trainX, trainPcData[,1], intercept=TRUE, lambda=mycv$lambda.min); rm(mycv)

lassoPredResult = predict(logitLasso, testX, type='response')
lassoPredResult = lassoPredResult>0.5

table(testPcData[,1]==1, lassoPredResult, dnn=c('Actual','Predicted'))

save(logitLasso, file='data/3-shrinkage/logitLasso.Rdata')
```







